---
title: "Assignment 3 R"
author: "Nabila Hossain"
date: "July 19, 2015"
output: 
  html_document: 
    keep_md: yes
---

I downloaded the mushroom dataset from https://archive.ics.uci.edu/ml/datasets/Mushroom. Then I uploaded the file to my github repository. I used the ```curl``` package, to retrieve the dataset from my repository.


```{r}
library(curl)
y <- read.csv(curl("https://raw.githubusercontent.com/nabilahossain/Image/master/assignment%203.txt"), header = FALSE)
head(y[1:10], 10)

```


I choose to look at the columns that hold information for "classes", "cap shape", "cap surface", "odor", "population" and "habitat" from the mushroom dataset. I choose appropriate columns and assign their names.

```{r}
x <- data.frame(y[c(1, 2, 3, 6, 22, 23)])
names(x) <- c("Classes", "Cap_Shape", "Cap_Surface", "Odor", "Population", "Habitat")
head(x, 3)
```

I check if I got the right number of rows and columns.
```{r}
ncol(x)
nrow(x)
```

I will replace the abbreviations for "classes", and check if they are correct.

```{r}
x$Classes <- ifelse(x$Classes=="p", "Poisonous", 
              ifelse(x$Classes=="e", "Edible", "N/A")
)
head(x, 5)
```

Next I will replace the abbreviations for columns: "classes", "cap_shape", "cap_surface", "odor", "population" and "habitat".

For Cap_Shape:
```{r}
x$Cap_Shape <- ifelse(x$Cap_Shape=="b", "Bell", 
              ifelse(x$Cap_Shape=="c", "Conical",
              ifelse(x$Cap_Shape=="x", "Convex",
              ifelse(x$Cap_Shape=="f", "Flat",
              ifelse(x$Cap_Shape=="k", "Knobbed",
              ifelse(x$Cap_Shape=="s", "Sunken", "N/A")
)))))
head(x, 3)
```        
                        
For Cap_Surface and Odor:
```{r}
x$Cap_Surface <- ifelse(x$Cap_Surface=="f", "Fibrous",
              ifelse(x$Cap_Surface=="g", "Grooves",
              ifelse(x$Cap_Surface=="y", "Scaly",
              ifelse(x$Cap_Surface=="s", "Smooth", "N/A")
)))
x$Odor <- ifelse(x$Odor=="a", "Almond",  
              ifelse(x$Odor=="l", "Anise",
              ifelse(x$Odor=="c", "Creosote",
              ifelse(x$Odor=="y", "Fishy",
              ifelse(x$Odor=="f", "Foul", 
              ifelse(x$Odor=="m", "Musty",
              ifelse(x$Odor=="n", "None",
              ifelse(x$Odor=="p", "Pungent",
              ifelse(x$Odor=="s", "Spicy", "N/A")
))))))))
tail(x, 3)
```                               

For Population and Habitat:
```{r}
x$Population <- ifelse(x$Population=="a", "Abundant",
              ifelse(x$Population=="c", "Clustered",
              ifelse(x$Population=="n", "Numerous",
              ifelse(x$Population=="s", "Scattered",
              ifelse(x$Population=="v", "Several",
              ifelse(x$Population=="y", "Solitary", "N/A")
)))))

x$Habitat <- ifelse(x$Habitat=="g", "grasses",  
              ifelse(x$Habitat=="l", "Leaves",
              ifelse(x$Habitat=="m", "Meadows",
              ifelse(x$Habitat=="p", "Paths",       
              ifelse(x$Habitat=="u", "Urban",
              ifelse(x$Habitat=="w", "Waste",
              ifelse(x$Habitat=="d", "Woods", "N/A")
))))))

head(x, 3) 
```


I check if I was able to replace the abbreviations for the subset correctly. I look at the first 30 rows (without unique row number) of the subset of mushrooms in a table format (using the ```knitr``` package):

```{r, tidy=TRUE}
rownames(x) <- NULL
library(knitr)
kable(head(x, 30))
```

If we wanted to look at only the first 15 edible mushrooms (without unique row number), in a table format, it would look like:
```{r, echo=TRUE, results="hold"} 
SetE <- subset(x, Classes=="Edible")
```

```{r, echo=FALSE, tidy=TRUE}
rownames(SetE) <- NULL
library(knitr)
kable(head(SetE, 15))
```